Bayesian Sparsification of $\cplx$-valued networks

### 2019-11-12 paper/paper.tex
finished compressing the appendix on the derivative

@@

\begin{equation}  \label{eq:g-m-deriv-int-1}
  \sum_{n\geq 0} \int_0^x \frac{t^{n+\alpha-1}}{n!} dt
    % = \int_0^x \sum_{n\geq 0} \frac{t^{n+\alpha-1}}{n!} dt
    = \int_0^x t^{\alpha-1} e^t dt
  % the MCT ([0, x], \mathcal{B}([0, x]), dx)
by the monotone convergence theorem, whence

@@

The substitution $u^2 = t$ on $[0, \infty]$ with $2u du = dt$ yields

@@

In particular, for $m=1$ we get
\begin{equation}  \label{eq:g-m-series-one}
  g_1(x)
    = - \gamma - 2 \log{2}
    + e^{-x} \sum_{n\geq 0} \frac{x^n}{n!} \sum_{p=1}^n \frac2{2p - 1}
and the derivative \eqref{eq:g-m-deriv} can be computed thus

@@

    = 2 \tfrac{F(\sqrt{x})}{\sqrt{x}}

@@

The derivative of \eqref{eq:ard-kl-div-real} with respect to $\alpha$ is
$$
\tfrac{d}{d \alpha} g_1(\tfrac1{2\alpha})
  = -\tfrac1{2 \alpha^2} g_1'(\tfrac1{2\alpha})
  % = -\tfrac1{2 \alpha^2} 2 \tfrac{F(\sqrt{\tfrac1{2\alpha}})}{\sqrt{\tfrac1{2\alpha}}}
  % = -\tfrac1{2 \alpha^2} 2 \tfrac{F(\tfrac1{\sqrt{2\alpha}})}{\tfrac1{\sqrt{2\alpha}}}
  = -\tfrac1{\alpha} \sqrt{\tfrac2{\alpha}} F(\tfrac1{\sqrt{2\alpha}})
  \,. $$
Since $\alpha$ is nonnegative, it is typically parameterized via its logarithm and computed
when needed. Thus the gradient of \eqref{eq:ard-kl-div-real} w.r.t $\log \alpha$ is
\begin{equation}  \label{eq:real-kl-div-deriv-log}
  \frac{d \eqref{eq:ard-kl-div-real}}{d\log \alpha}
  % \frac{d}{d\log \alpha}
  %   \frac12 \mathbb{E}_{z \sim \mathcal{N}(0,1)}
  %     \log \lvert z + \tfrac1{\sqrt{\alpha}} \rvert^2
    % = \frac12 \frac{d\alpha}{d\log \alpha}
    %   \frac{d}{d\alpha} \bigl(\mathbb{E}\cdots \bigr)
    % = - \frac\alpha2 \tfrac1{\alpha}
    %   \sqrt{\tfrac2{\alpha}} F(\tfrac1{\sqrt{2\alpha}})
    = - \frac1{\sqrt{2\alpha}} F\bigl(\tfrac1{\sqrt{2\alpha}}\bigr)
    \,.
\end{equation}

We replicated the simulation experiment in \citep{molchanov_variational_2017} and recovered
the same nonlinear approximation result reported therein. We sampled $10^7$ draws of  and 
The coefficients are within $1\%$ relative tolerance of the reported.
$$
  \eqref{eq:ard-kl-div-real}
% - KL(\mathcal{N}(w \mid \mu, \alpha \mu^2) \|
%     \tfrac1{\lvert w \rvert})
%   \approx
%     k_1 \sigma(k_2 + k_3 \log \alpha)
%     + C \big\vert_{C = -k1}
%     - \tfrac12 \log (1 + e^{-\log \alpha})
  % = \tfrac12 \log \alpha
  %   - \mathbb{E}_{\xi \sim \mathcal{N}(1, \alpha)}
  %   \log{\lvert \xi \rvert} + C
  \approx
    \frac12 \log{\bigl(1 + e^{-\log \alpha}\bigr)}
    + k_1 \sigma\bigl(- (k_2 + k_3 \log \alpha)\bigr)
    % - k_1 \bigl(1 - \sigma(k_2 + k_3 \log \alpha)\bigr)
  \,. $$
% in parctial implementations we suggest using softplus to numerically comptue the last
% term, since it is typically implemented in a floating-point stable fashion.

@@

Now in SGD (and specifically in SGVB) we are concerned more with the gradient field, induced
by the potential (which is the loss function), rather than the value itself. Thus, as far as
the regularizing penalty term is concerned, which is used to regularize the loss objective,
we can essentially ignore it's forward pass value (level) and just compute its gradient
(subgradient, normal cone) with respect the parameter of interest in sgd. (unless it is a
part of a constraint, ie. downstream computation)

The numerically estimated derivative of the penalty term wrt $\log \alpha$ using forward
differences seems to be following \eqref{eq:real-kl-div-deriv-log} very closely (up to
sampling error) see fig.~\ref{fig:molchanov-derivative-replica}.

>>>>

% In stochastic gradient methods are concerned more with the gradient, induced by the objective
% function, rather than its value.

@@

% by the monotone convergence theorem on ([0, x], \mathcal{B}([0, x]), dx)

@@

Substitution $u^2 = t$ on $[0, \infty]$ with $2u du = dt$ yields

@@

In particular, the derivative of \eqref{eq:g-m-series} for $m=1$ is

@@

    = 2 \frac{F(\sqrt{x})}{\sqrt{x}}

@@

Hence, the derivative of \eqref{eq:ard-kl-div-real} with respect to $\alpha$ is
$$
\frac{d}{d \alpha} g_1(\tfrac1{2\alpha})
  = -\frac1{2 \alpha^2} g_1'(\tfrac1{2\alpha})
  % = -\frac1{2 \alpha^2} 2 \tfrac{F(\sqrt{\tfrac1{2\alpha}})}{\sqrt{\tfrac1{2\alpha}}}
  % = -\frac1{2 \alpha^2} 2 \tfrac{F(\tfrac1{\sqrt{2\alpha}})}{\tfrac1{\sqrt{2\alpha}}}
  = -\frac1{\alpha} \sqrt{\frac2{\alpha}} F(\tfrac1{\sqrt{2\alpha}})
  \,. $$
Finally, since $\alpha$ is nonnegative, it is typically parameterized via its logarithm.
The gradient of \eqref{eq:ard-kl-div-real} w.r.t $\log \alpha$ is
\begin{equation}  \label{eq:real-kl-div-deriv-log}
  \frac{d \eqref{eq:ard-kl-div-real}}{d\log \alpha}
  % \frac{d}{d\log \alpha}
  %   \frac12 \mathbb{E}_{z \sim \mathcal{N}(0,1)}
  %     \log \lvert z + \tfrac1{\sqrt{\alpha}} \rvert^2
    % = \frac12 \frac{d\alpha}{d\log \alpha}
    %   \frac{d}{d\alpha} \bigl(\mathbb{E}\cdots \bigr)
    % = - \frac\alpha2 \tfrac1{\alpha}
    %   \sqrt{\tfrac2{\alpha}} F(\tfrac1{\sqrt{2\alpha}})
    = - \frac1{\sqrt{2\alpha}} F\bigl(\tfrac1{\sqrt{2\alpha}}\bigr)

We compute the Monte-Carlo estimator of \eqref{eq:ard-kl-div-real} on a sample of $10^7$
draws over an equally spaced grid of $\log \alpha$ in $[-12, +12]$ of size $4096$. The optimal
coefficients of the regression, proposed by \citep{molchanov_variational_2017}, are within
$1\%$ relative tolerance of the reported therein.
\begin{equation}  \label{eq:ard-kl-div-real-approx}
  \eqref{eq:ard-kl-div-real}
% - KL(\mathcal{N}(w \mid \mu, \alpha \mu^2) \|
%     \tfrac1{\lvert w \rvert})
%   \approx
%     k_1 \sigma(k_2 + k_3 \log \alpha)
%     + C \big\vert_{C = -k1}
%     - \tfrac12 \log (1 + e^{-\log \alpha})
  % = \tfrac12 \log \alpha
  %   - \mathbb{E}_{\xi \sim \mathcal{N}(1, \alpha)}
  %   \log{\lvert \xi \rvert} + C
  \approx
    \frac12 \log{\bigl(1 + e^{-\log \alpha}\bigr)}
    + k_1 \sigma\bigl(- (k_2 + k_3 \log \alpha)\bigr)
    % - k_1 \bigl(1 - \sigma(k_2 + k_3 \log \alpha)\bigr)
  \,,
\end{equation}
% in parctial implementations we suggest using softplus to numerically comptue the last
% term, since it is typically implemented in a floating-point stable fashion.
The numerically estimated derivative of the penalty term wrt $\log \alpha$ using forward
differences seems to be following \eqref{eq:real-kl-div-deriv-log} very closely (up to
sampling error) see fig.~\ref{fig:molchanov-derivative-replica}. The derivative of
\eqref{eq:ard-kl-div-real-approx} with respect to $\log \alpha$ appears to be very close
to \eqref{eq:real-kl-div-deriv-log}.

@@
